The Error Estimate of Force Calculation in the Inhomogeneous Molecular Systems

### 2012-04-23 shortRange.inhomo/alltext/pubilsh/main.tex
add missing files
trying to make everything better.

\documentclass[aps,pre,preprint]{revtex4}
% \documentclass[aps,pre,twocolumn]{revtex4-1}
% \documentclass[aps,jcp,groupedaddress,twocolumn,unsortedaddress]{revtex4}

@@

Fortunately,
in most homogeneous systems, a satisfactory convergence can be
achieved even with a very small cut-off value, for example,
$2.5\sigma$ for the standard Lennard-Jones 6-12 interaction.
Moreover, it is possible to correct the systematic error
of the potential energy, the pressure and the free energy by applying
the standard long range correction (LRC)~\cite{allen87a}, which
assumes the uniformity outside the cut-off radius and integrates the
contribution of the mentioned properties to infinity. 
% The spacial inhomogeneity of the system in terms of particle density,
% particle type and change distribution is very common in natrual
% system.
Instead for inhomogeneous systems, in which the density of a certain type of
atom may change within the length scale of a few molecules, 
and thus lack of uniformity, 
the standard molecular simulation 
by the cut-off method will encounter non-trivial
difficulties.
The simulation results are found to converge only at a very large
cut-off radius value~\cite{duque2004some, ismail2007application}, also
depend on the energy/force continuity at the cut-off and the way of
applying the long-range corrections~\cite{trokhymchuk1999computer,
  guo1997long, mecke1997molecular, janecek2006long, goujon2004monte,
  shen2007comparative}.  Therefore, it is necessary to systematically
study the precision of the short-range interaction calculation, and to
develop methods for screening out the aforementioned cut-off artifact in
the inhomogeneous systems.

% In this paper, we focus on the systems in which
% atoms/particles are interacting via short-range interactions, because
% it is a good starting point of the study of more complex systems, for
% example, the systems with long-range electrostatic interactions.  The
% simulation results in the inhomogeneous systems are found to depend on
% a number of factors of treating the short-range interaction in the
% simulation. For example, the way of calculating the short-range
% interaction (simple cut-off method or Ewald
% summation)~\cite{karasawa1989acceleration, lopez2002thermodynamic,
%   lopez2006liquid, shi2006molecular, ismail2007application}, the
% cut-off radius~\cite{trokhymchuk1999computer, duque2004some,
%   shen2007comparative}, the way of long-range
% correction~\cite{guo1997long, mecke1997molecular, janecek2006long,
%   goujon2004monte} and the size and shape of the simulation
% box~\cite{chen1995area, orea2005oscillatory, biscay2009calculation}.
% It should be noticed that the physical properties are the nature of
% the system, and should NOT depend on the details of the interaction
% calculation.

A promising way to quantitatively analyze the undesirable cut-off
effects is to express these artifacts in terms of the difference
between the cut-offed interaction and the exact interaction, namely
the error. In homogeneous systems, the error analysis of the cut-off
method has been well established, see~\cite{kolafa1992cutoff, wang2011} for
example. Also the error estimates~\cite{hummer1995numerical,
  kolafa1992cutoff, petersen1995accuracy, deserno1998mue2,
  wang2010optimizing} for long-range Ewald family
methods~\cite{ewald1921die, hockney1988computer, deserno1998mue1,
  darden1993pme, essmann1995spm} have provided a profound
understanding of the accuracy of force calculation, and introduced
parameter tuning algorithms~\cite{limbach06a, wang2010optimizing} that
boost the efficiency of the computation with a good control of the
error. However, in inhomogeneous systems the error study is still
scarce, even for the simplest cut-off method.

In this work, we develop the error estimate of the short-range force
in inhomogeneous systems. We prove that the 
force error is comprised of a homogeneity error and an inhomogeneity
error, regardless of the particle-particle correlation outside the
cut-off radius.  In an inhomogeneous system composed by several bulk
regions connected by interfacial regions where the density changes
rapidly, an important observation is that the
inhomogeneity error plays a dominant role in the interfacial regions,
and can be more than one order of magnitude larger than
the error in the bulk regions. An
adaptive cut-off method is hereby proposed to equally distribute the
error across the simulation region, so that a considerable
computational effort can be saved in the bulk regions
while the precision of simulation is
preserved. We also proposed a long-range force correction (LFC) method
to screen out the contribution of the
inhomogeneity error, therefore,  the precision of
simulation is improved without much extra computational load. To show the
effectiveness of the error estimate and the proposed methods, two
systems are tested: the liquid-vapor equilibrium for equilibrium
properties, and the colliding nanoscale clusters for dynamical
properties. Finally, conclusions are given in section
\ref{sec:conclusion}.

@@

In this paper, for simplicity,
we will consider one-component systems, but
% we only consider the single-component system. Because
% this assumption simplifies the discussion, and
the error estimates for
the one-component system can be straightforwardly extended to the
multi-component systems.  We consider
the force error in the present paper, because the
applications are based on molecular dynamics simulations.
% Since the error estimate developed for the one-component system can be
% directly extended to multi-component system without substantial
% difficulty, the discussion in this paper will focus on one-component
% system for symplicity. 
Suppose the system is composed by $N$ identical particles located at
$\v r_1, \v r_2, \cdots, \v r_N$ with periodic boundary condition.
These particles are interacting via a short-range pairwise interaction
$u(\v r)$ with a cut-off radius $r_c > 0$. The corresponding force is
denoted by $\v f(\v r)$. 
% An infinite cut-off radius will lead to
% a precise calculation of interaction, but it is impractical.
% The error of computing interaction is mainly introduced by a finit
% value of the cut-off radius.
The force error mainly stems from the neglected interaction outside
the cut-off radius, so it is convenient to consider the complementary
of the cut-offed force that is defined by
\begin{align}
  \v f^c(\v r) =
  \left\{
  \begin{array}{ll}
    0, & \vert\v r\vert\leq r_c; \\
    \v f(\v r), & \vert\v r\vert > r_c.
  \end{array}
  \right.
\end{align}
We measure the force error at an arbitrary position in the system,
saying $\v r$, by adding a testing particle to that position.  We
assume all particles in the system exert forces on the testing
particle, whereas the testing particle does NOT exert any force on the
system.  The exact force and the computed force (or cut-offed force)
exerted by the system on the testing particle are denoted by ${\v
  F}(\v r)$ and $\tilde{\v F}(\v r)$, respectively.  The difference
between them, namely the \emph{error force}, is denoted by $\Delta\v
F(\v r) = \v F(\v r) - \tilde{\v F}(\v r)$, which can be expressed by
\begin{align}
  \Delta\v F(\v r) = \sum_{\v n}\sum_{j}\, \v f^c(\v r - \v r_j + \v n).
\end{align}
Where $\v n$ is the box vector, so the summation over $\v n$ indicates
that all periodic images of the system are considered. In a
homogeneous system, the surrounding of any particle is isotropic, so
the \emph{mean error force} vanishes: $\langle\Delta\v F(\v r)\rangle
= 0$. In contrast, the mean error force is not necessarily zero in an
inhomogeneous system:
\begin{align} \label{eqn:tmp3} %\nonumber
  \langle\Delta\v F(\v r)\rangle
  &=
  \big\langle
  \sum_{\v n}\sum_{j}\, \v f^c(\v r - \v r_j + \v n)
  \big\rangle 
  =
  \int_{\mathbb R^3}\v f^c(\v r - \v r')\rho(\v r')\,\d d\v r',
\end{align}
where $\rho(\v r)$ denotes the particle number density at the position
$\v r$, and is periodically extended to $\mathbb R^3$. Notice that the
error force implicitly depends on the cut-off radius used for the
interaction calculation.

@@

\section*{Acknowledgments}
The authors gratefully thank Luigi Delle Site for for useful comments
on the manuscript. The work of P.Z. is supported by grants from
National Natural Science Foundation of China (50930003).

>>>>

\documentclass[preprint]{revtex4}
% \documentclass[aps,pre,twocolumn]{revtex4-1}
% \documentclass[aps,jcp,groupedaddress,twocolumn,unsortedaddress]{revtex4}

@@

Fortunately,
in most homogeneous systems, a satisfactory convergence can be
achieved even with a very small cut-off value, for example,
$2.5\sigma$ for the standard Lennard-Jones 6-12 interaction.
Moreover, it is possible to correct the systematic error
of the potential energy, the pressure and the free energy by applying
the standard long-range correction (LRC)~\cite{allen87a}, which
assumes the uniformity outside the cut-off radius and integrates the
contribution of the mentioned properties to infinity. 
% The spacial inhomogeneity of the system in terms of particle density,
% particle type and change distribution is very common in natrual
% system.
Instead for inhomogeneous systems, in which the density of a certain type of
atom may change within the length scale of a few molecules, 
and thus lack of uniformity, 
the standard molecular simulation 
by the cut-off method will encounter non-trivial
difficulties.
The simulation results are found to converge only at a very large
cut-off radius value~\cite{duque2004some, ismail2007application}, also
depend on the energy/force continuity at the cut-off and the way of
applying the long-range corrections~\cite{trokhymchuk1999computer,
  guo1997long, mecke1997molecular, janecek2006long, goujon2004monte,
  shen2007comparative}.  Therefore, it is necessary to systematically
study the precision of the short-range interaction calculation, and to
develop methods for screening out the aforementioned cut-off artifact in
the inhomogeneous systems.

\recheck{ One possible way to avoid the cut-off difficulty is to
  modify the short-range interaction by a shifting approach: inside a
  certain radius the origianl interaction is kept, while outside it the
  interaction is screen by a shift function, so that goes to
  zero smoothly. The amount of Shifting is minor in most cases,
  however, it actually changes the way of
  modelling molecules. For example, it
  has been shown that shifting the dispersion remarkably changes the
  phase diagram of the Lennard-Jones system~\cite{smit1992phase}.
  Since the dispersion term 
  can be derived by quantum mechanical arguments, and has the physical
  reality, the unshifted interaction is
  prefered and investigated by most studies
  (see e.g. Ref.~\cite{potoff1998critical, perez2006critical, wang2010optimizing}).
  % Shifting the direct part of the Ewald summation actually
  % violates the Coulomb's law in the way that depends on the
  % splitting parameter, so it should be treated carefully.
  % Here we would like to suggest treating the shifting approach
  % as a special case 
}
% In this paper, we focus on the systems in which
% atoms/particles are interacting via short-range interactions, because
% it is a good starting point of the study of more complex systems, for
% example, the systems with long-range electrostatic interactions.  The
% simulation results in the inhomogeneous systems are found to depend on
% a number of factors of treating the short-range interaction in the
% simulation. For example, the way of calculating the short-range
% interaction (simple cut-off method or Ewald
% summation)~\cite{karasawa1989acceleration, lopez2002thermodynamic,
%   lopez2006liquid, shi2006molecular, ismail2007application}, the
% cut-off radius~\cite{trokhymchuk1999computer, duque2004some,
%   shen2007comparative}, the way of long-range
% correction~\cite{guo1997long, mecke1997molecular, janecek2006long,
%   goujon2004monte} and the size and shape of the simulation
% box~\cite{chen1995area, orea2005oscillatory, biscay2009calculation}.
% It should be noticed that the physical properties are the nature of
% the system, and should NOT depend on the details of the interaction
% calculation.

A promising way to quantitatively analyze the undesirable cut-off
effects is to express these artifacts in terms of the difference
between the cut-offed interaction and the exact interaction, namely
the error. In homogeneous systems, the error analysis of the cut-off
method has been well established, see~\cite{kolafa1992cutoff, wang2011} for
example. Also the error estimates~\cite{hummer1995numerical,
  kolafa1992cutoff, petersen1995accuracy, deserno1998mue2,
  wang2010optimizing} for long-range Ewald family
methods~\cite{ewald1921die, hockney1988computer, 
  darden1993pme, essmann1995spm, deserno1998mue1} have provided a profound
understanding of the accuracy of force calculation, and introduced
parameter tuning algorithms~\cite{limbach06a, wang2010optimizing} that
boost the efficiency of the computation with a good control of the
error. However, in inhomogeneous systems the error study is still
scarce, even for the simplest cut-off method.

In this work, we develop the error estimate of the short-range force
in inhomogeneous systems. We prove that the 
force error is comprised of a homogeneity error and an inhomogeneity
error, regardless of the particle-particle correlation outside the
cut-off radius.  In an inhomogeneous system composed by several bulk
regions connected by interfacial regions where the density changes
rapidly, an important observation is that the
inhomogeneity error plays a dominant role in the interfacial regions,
and can be more than one order of magnitude larger than
the error in the bulk regions. An
adaptive cut-off method is hereby proposed to equally distribute the
error across the simulation region, so that a considerable
computational effort can be saved in the bulk regions
while the precision of simulation is
preserved. We also proposed a long-range force correction (LFC) method
to screen out the contribution of the
inhomogeneity error, therefore,  the precision of
simulation is improved without much extra computational load.

To show the
effectiveness of the error estimate and the proposed methods, two
systems are tested: the liquid-vapor equilibrium for equilibrium
properties, and the colliding nanoscale clusters for dynamical
properties.
\recheck{
  In these example, we investigate the Lennard-Jones interaction, which
  is one of the most widely used in molecular simulations.
  However, the error estimate,
  the adaptive cut-off and long-range correction methods
  are not restricted to the Lennard-Jones: it can
  be applied to other types of short-range interactions without 
  difficulty, such as
  Buckingham, the direct part of Ewald summation~\cite{ewald1921die}, etc..
  We want to stress that the main contribution and the research interest is the
  quantitative study of the force computation error by cuting-off
  the short-range interaction.
  The examples, which are comparatively simple and ideal,
  are used only for testing the validity of the error
  estimate and the proposed methods,
  and demostrating the role of different errors in the simulations.
  The error study in more realistic systems requires the
  estimates of the electrostatic interaction,
  which are calculated by the long-range algorithms, such as
  PME~\cite{darden1993pme}, SPME~\cite{essmann1995spm},
  P3M~\cite{deserno1998mue1}, etc..
  The space of the present paper is not enough for
  these algorithms, so they will be delivered in the future work.
}
% Finally, conclusions are given in section
% \ref{sec:conclusion}.

@@

In this paper, for simplicity,
we will consider one-component systems, but
% we only consider the single-component system. Because
% this assumption simplifies the discussion, and
the error estimate for
the one-component system can be straightforwardly extended to the
multi-component systems.  We consider
the force error in the present paper, because the
applications are based on molecular dynamics simulations.
% Since the error estimate developed for the one-component system can be
% directly extended to multi-component system without substantial
% difficulty, the discussion in this paper will focus on one-component
% system for symplicity. 
Suppose the system is composed by $N$ identical particles located at
$\v r_1, \v r_2, \cdots, \v r_N$ with periodic boundary condition.
These particles are interacting via a short-range pairwise interaction
$u(\v r)$ with a cut-off radius $r_c > 0$. The corresponding force is
denoted by $\v f(\v r)$. 
% An infinite cut-off radius will lead to
% a precise calculation of interaction, but it is impractical.
% The error of computing interaction is mainly introduced by a finit
% value of the cut-off radius.
The force error mainly stems from the neglected interaction outside
the cut-off radius, so it is convenient to consider the complementary
of the cut-offed force that is defined by
\begin{align}
  \v f^c(\v r) =
  \left\{
  \begin{array}{ll}
    0, & \vert\v r\vert\leq r_c; \\
    \v f(\v r), & \vert\v r\vert > r_c.
  \end{array}
  \right.
\end{align}
We measure the force error at an arbitrary position in the system,
saying $\v r$, by adding a testing particle to that position.  We
assume all particles in the system exert forces on the testing
particle, whereas the testing particle does NOT exert any force on the
system.  The exact force and the computed force (or cut-offed force)
exerted by the system on the testing particle are denoted by ${\v
  F}(\v r)$ and $\tilde{\v F}(\v r)$, respectively.  The difference
between them, namely the \emph{error force}, is denoted by $\Delta\v
F(\v r) = \v F(\v r) - \tilde{\v F}(\v r)$, which can be expressed by
\begin{align}
  \Delta\v F(\v r) = \sum_{\v n}\sum_{j}\, \v f^c(\v r - \v r_j + \v n).
\end{align}
Where $\v n$ is the box vector, so the summation over $\v n$ indicates
that all periodic images of the system are considered. In a
homogeneous system, the surrounding of any particle is isotropic, so
the \emph{mean error force} vanishes: $\langle\Delta\v F(\v r)\rangle
= 0$. In contrast, the mean error force is not necessarily zero in an
inhomogeneous system:
\begin{align} \label{eqn:tmp3} %\nonumber
  \langle\Delta\v F(\v r)\rangle
  &=
  \big\langle
  \sum_{\v n}\sum_{j}\, \v f^c(\v r - \v r_j + \v n)
  \big\rangle 
  =
  \int_{\mathbb R^3}\v f^c(\v r - \v r')\rho(\v r')\,\d d\v r',
\end{align}
where $\rho(\v r)$ denotes the particle number density at the position
$\v r$, and is periodically extended to $\mathbb R^3$. Notice that the
error force implicitly depends on the cut-off radius used for the
interaction calculation.

@@

\recheck{
  We want to stress again the main research interest of the present paper is the
  error estimate, the proposed adaptive cut-off and long-range
  correction methods. The examples are used to demonstrate
  the validity of these theoretical results. The applications
  of the error estimate can be more general than the studied examples.
  Other kind the short-range interaction (not only the Lennard-Jones
  but also the Buckingham,
  the direct part of the Ewald summation, etc.),
  and other kinds of inhomogeneity
  (not only  planar and merging spirical interfaces)
  can be easily studied by the error estiamtes and the proposed methods.
  The error study of
  more realistic systems requires the error estiamtes of the long-range
  algoritms, and careful comparisons with the short-range error.
  Since the present paper is focusing on the short-range
  interactions, and the space is limited,
  they will be given in the future work.
}

\section*{Acknowledgments}
The authors gratefully thank Luigi Delle Site for useful comments
on the manuscript. The work of P.Z. is supported by grants from
National Natural Science Foundation of China (50930003).
